### 2023-03-27 main/chapter_experiment.tex
add fcc etc writing
add breakdown of lumi/conditions by year

\section{Inner detector}
\label{sec:atlas_inner}
The inner tracking detectors begin \atlas's measurements of new collisions
by precisely tracing the paths taken by energetic charged particles
as they curve through the solenoidal magnetic field.
Track curvature provides momentum information,
tracks' associations into collision vertices helps to separate the multiple
scatterings per crossing (pile-up),
and their associations into displaced vertices help to identify
decays.
\begin{quote}
``%
The inner tracking detector covers the pseudorapidity range $|\eta| < 2.5$.
It consists of silicon pixel, silicon microstrip, and transition radiation
tracking detectors.
An additional layer of silicon pixels, the insertable
B-layer~\cite{ATLAS-TDR-19, PIX-2018-001}, was installed before Run~2.%
''~\cite{atlas2022searches}
\end{quote}
This insertable B-layer is a silicon pixel detector that begins just
$3\,\textrm{cm}$ from the beam-pipe, and provides robust tracking for the
high luminosity of the current LHC,%
\footnote{%
Instantaneous luminosity in the Run~2 LHC peaked at
$19\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$ in 2018~\cite{ATLAS:2022hro},
which is much higher than previous conditions such as the Run~1 peak
instantaneous luminosity of
$3.6\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$ in 2012~\cite{ATLAS:2013tpw}.
The future High Luminosity LHC (HL-LHC), however, might achieve
$750\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$.
For the HL-LHC, the current inner tracking and insertable B-layer will be
replaced by a new Inner Trac\emph{k}er (ITk) which is designed to handle such
extreme luminosity conditions~\cite{CERN-LHCC-2017-021}.
}
particularly for sensitivity to particles whose lifetimes are just right to
decay inside the tracking system;
these `Goldilocks Zone' particles include B-hadrons, some tau leptons, and
maybe new species.

@@

The search of this thesis uses data from proton-proton collisions at
$\sqrt{s} = 13\,\eV[T]$ centre-of-mass energies in Run~2 of the
LHC, in which \atlas\ collected data in the years
$2015\textrm{--}2018$~\cite{atlas2022searches}.
This second run has various refinements over the Run~1, including beam energies
increasing from $\sqrt{s} = 7\textrm{ or }8\,\eV[T]$, and bunch spacings
halving from $50$ to $25\,\textrm{ns}$~\cite{lhc2006run2}.

Data collected under anomalous detector operations are hard to model.
Lists are maintained indicating which datasets arise from `good', normal
detector operations, and these lists are commonly employed to select only clean
subsets of good data for analysis projects.
Testament to the quality of \atlas's operations, this cleaning removes less
than $5\%$ of the Run~2 data~\cite{
DAPR-2018-01,
Golling:2011zy
}.

% closing comment?
Now that we have predictive models in the Standard Model and supersymmetric
alternatives
and interpretable data from \atlas's measurements of LHC collisions,
we are ready to begin to squeeze some useful research from those data.

>>>>

\section{Inner detector}
\label{sec:atlas_inner}
The inner tracking detectors begin \atlas's measurements of new collisions
by precisely tracing the paths taken by energetic charged particles
as they curve through the solenoidal magnetic field.
Track curvature provides momentum information,
tracks' associations into collision vertices helps to separate the multiple
scatterings per crossing (pile-up),
and their associations into displaced vertices help to identify
decays.
\begin{quote}
``%
The inner tracking detector covers the pseudorapidity range $|\eta| < 2.5$.
It consists of silicon pixel, silicon microstrip, and transition radiation
tracking detectors.
An additional layer of silicon pixels, the insertable
B-layer~\cite{ATLAS-TDR-19, PIX-2018-001}, was installed before Run~2.%
''~\cite{atlas2022searches}
\end{quote}
This insertable B-layer is a silicon pixel detector that begins just
$3\,\textrm{cm}$ from the beam-pipe, and provides robust tracking for the
high luminosity of the current LHC,%
\footnote{%
Instantaneous luminosity in the Run~2 LHC peaked at
$19\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$ in 2018~\cite{ATLAS:2022hro},
which is much higher than previous conditions such as the Run~1 peak
instantaneous luminosity of
$3.6\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$ in 2012~\cite{ATLAS:2013tpw}.
The future High-Luminosity LHC (HL-LHC), however, might achieve
$75\times10^{33}\,\mathrm{cm}^{-2}\mathrm{s}^{-1}$.
For the HL-LHC, the current inner tracking and insertable B-layer will
be replaced by a new Inner Trac\emph{k}er (ITk) which is designed to handle such
extreme luminosity conditions~\cite{CERN-LHCC-2017-021}.
}
particularly for sensitivity to particles whose lifetimes are just right to
decay inside the tracking system;
these `Goldilocks Zone' particles include B-hadrons, some tau leptons, and
maybe new species.

@@

The search of this thesis uses data from proton-proton collisions at
$\sqrt{s} = 13\,\eV[T]$ centre-of-mass energies in Run~2 of the
LHC, in which \atlas\ collected data in the years
$2015\textrm{--}2018$~\cite{atlas2022searches}.
This second run has various refinements over the Run~1, including beam energies
increasing from $\sqrt{s} = 7\textrm{ or }8\,\eV[T]$, and bunch spacings
halving from $50$ to $25\,\textrm{ns}$~\cite{lhc2006run2}.
A breakdown of the LHC proton-proton data conditions and quantity is given
in Table~\ref{tab:experiment_run2_state}.
Broadly, instantaneous luminosity was increased over time, leading to increases
in both data quantity and pile-up interactions alongside those data.

\begin{sidewaystable}[tp]
\centering
\begin{tabular}{lcccc}
Observable                                                        & 2015         & 2016         & 2017                & 2018         \\
\hline
Maximum number of colliding bunch-pairs                           & $2232$       & $2208$       & $2544$ ($1909$)     & $2544$       \\
Bunch spacing ($\mathrm{ns}$)                                     & $25$         & $25$         & $25$ (---)          & $25$         \\
Typical number of protons per bunch ($\times 10^{11}$)            & $1.1$        & $1.1$        & $1.1$ ($1.2$)       & $1.1$        \\
Peak luminosity ($\times 10^{33}\mathrm{cm}^{-2}\mathrm{s}^{-1}$) & $5$          & $13$         & $16$                & $19$         \\
Peak number of inelastic interactions per crossing                & $\approx 16$ & $\approx 41$ & $\approx 45$ ($60$) & $\approx 55$ \\
Luminosity-weighted average inelastic interactions per crossing   & $13$         & $25$         & $38$                & $36$         \\
Integrated luminosity delivered ($\mathrm{fb}^{-1}$)              & $4$          & $39$         & $50.6$              & $63.8$
\end{tabular}
\caption[
LHC data rates and conditions by year in Run~2
]{%
LHC data rates and conditions by year in Run~2.
This table is adapted from~\cite{ATLAS:2022hro}.
Bracketed numbers under 2017 refer to an alternative LHC operating condition
which alternated in a pattern of eight bunches with the usual
$25\,\mathrm{ns}$ separation followed by a gap of four empty bunch slots.
Alongside a primary interaction, the inelastic interactions in a crossing are
pile-up events and add noise to our measurements.
}
\label{tab:experiment_run2_state}
\end{sidewaystable}

Data collected under anomalous detector operations are hard to model.
Lists are maintained indicating which datasets arise from `good', normal
detector operations, and these lists are commonly employed to select only clean
subsets of good data for analysis projects.
Testament to the quality of \atlas's operations, this cleaning removes less
than $5\%$ of the Run~2 data~\cite{
DAPR-2018-01,
Golling:2011zy
}.
After selecting for data quality requirements, we are left with a total of
$139.0\,\mathrm{fb}^{-1}$ \atlas\ proton-proton collision data to analyse.
This integrated luminosity comprises
% lumi_15_16 = 36.2 # [1/fb] 15+16
% lumi_17 = 44.3 # [1/fb] 17
% lumi_18 = 58.5 # [1/fb] 18
$36.2\,\mathrm{fb}^{-1}$ from the combination of 2015 and 2016,
$44.3\,\mathrm{fb}^{-1}$ from 2017, and
$58.5\,\mathrm{fb}^{-1}$ from 2018.

% closing comment
Now that we have predictive models in the Standard Model and supersymmetric
alternatives
and interpretable data from \atlas's measurements of LHC collisions,
we are ready to begin to squeeze some useful research from those data.
